Capstone Movielens Report

### 2025-05-05 reports/capstone-movielens-report.draft4.Rmd
Draft report updated.

##### 
\
```{r}

\
```{r echo=FALSE}
# #### [@MRS-R-BEST]
```

## Methods / Analysis
::: {.noteblock data-latex=""}
All the source code of the R-scripts is available on the project's [GitHub repository](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/tree/main/r/src). 
:::

@@

We will split the `edx` dataset into a training set, which we will use to build and train our models, and a test set in which we will compute the accuracy of our predictions, the way described in [Section 23.1.1 Movielens data](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movielens-data) of the _Course Textbook_ mentioned above[@IDS2_23-1-1].We will also use the _5-Fold Cross Validation_ method the way described in [Section 29.6 Cross validation](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#cross-validation) of the _Course Textbook_. 
To prepare datasets for processing, we will use the following functions, specifically designed for these operations:

- `make_source_datasets`;
- `init_source_datasets`.

@@

```{r eval=FALSE, echo=TRUE, include=TRUE}
# Let's ignore the data for users who have not provided at least 100 ratings:
edx100 <- edx |> 
  group_by(userId) |>
  filter(n() >= 100) |>
  ungroup()

print(edx100 |> summarize(n_distinct(userId), n_distinct(movieId)))

# For each one of these users, we will split their ratings into 80% for training 
# and 20% for testing:

set.seed(2006)
indexes <- split(1:nrow(edx100), edx100$userId)
test_ind <- sapply(indexes, function(i) sample(i, ceiling(length(i)*.2))) |> 
  unlist() |>
  sort()

test_set <- edx100[test_ind,] 
train_set <- edx100[-test_ind,]

# To make sure we don’t include movies in the training set that should not be 
# there, we remove entries using the semi_join function:
test_set <- test_set |> semi_join(train_set, by = "movieId") |> as.data.frame()
summary(test_set)

train_set <- mutate(train_set, userId = factor(userId), movieId = factor(movieId))
summary(train_set)
```

>>>>

## Methods / Analysis
::: {.noteblock data-latex=""}
All the source code of the R-scripts is available on the project's [GitHub repository](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/tree/main/r/src)[@edx_capstone_movielens]. 
:::

@@

We will split the `edx` dataset into a training set, which we will use to build and train our models, and a test set in which we will compute the accuracy of our predictions, the way described in [Section 23.1.1 Movielens data](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movielens-data) of the _Course Textbook_ mentioned above[@IDS2_23-1-1].We will also use the _5-Fold Cross Validation_ method the way described in [Section 29.6 Cross validation](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#cross-validation) of the _Course Textbook_. 
To prepare datasets for processing, we will use the following functions, specifically designed for these operations:
```{r eval=FALSE }
make_source_datasets()
init_source_datasets()
```
- `make_source_datasets`;
- `init_source_datasets`.

@@

### 2025-05-05 reports/capstone-movielens-report.draft4.tex
Draft report updated.

\subparagraph{}\label{section}

@@

\begin{noteblock}
All the source code of the R-scripts is available on the project's
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/tree/main/r/src}{GitHub
repository}.

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Let\textquotesingle{}s ignore the data for users who have not provided at least 100 ratings:}
\NormalTok{edx100 }\OtherTok{\textless{}{-}}\NormalTok{ edx }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{group\_by}\NormalTok{(userId) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{n}\NormalTok{() }\SpecialCharTok{\textgreater{}=} \DecValTok{100}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{ungroup}\NormalTok{()}

\FunctionTok{print}\NormalTok{(edx100 }\SpecialCharTok{|\textgreater{}} \FunctionTok{summarize}\NormalTok{(}\FunctionTok{n\_distinct}\NormalTok{(userId), }\FunctionTok{n\_distinct}\NormalTok{(movieId)))}

\CommentTok{\# For each one of these users, we will split their ratings into 80\% for training }
\CommentTok{\# and 20\% for testing:}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2006}\NormalTok{)}
\NormalTok{indexes }\OtherTok{\textless{}{-}} \FunctionTok{split}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{nrow}\NormalTok{(edx100), edx100}\SpecialCharTok{$}\NormalTok{userId)}
\NormalTok{test\_ind }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(indexes, }\ControlFlowTok{function}\NormalTok{(i) }\FunctionTok{sample}\NormalTok{(i, }\FunctionTok{ceiling}\NormalTok{(}\FunctionTok{length}\NormalTok{(i)}\SpecialCharTok{*}\NormalTok{.}\DecValTok{2}\NormalTok{))) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{unlist}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{sort}\NormalTok{()}

\NormalTok{test\_set }\OtherTok{\textless{}{-}}\NormalTok{ edx100[test\_ind,] }
\NormalTok{train\_set }\OtherTok{\textless{}{-}}\NormalTok{ edx100[}\SpecialCharTok{{-}}\NormalTok{test\_ind,]}

\CommentTok{\# To make sure we don’t include movies in the training set that should not be }
\CommentTok{\# there, we remove entries using the semi\_join function:}
\NormalTok{test\_set }\OtherTok{\textless{}{-}}\NormalTok{ test\_set }\SpecialCharTok{|\textgreater{}} \FunctionTok{semi\_join}\NormalTok{(train\_set, }\AttributeTok{by =} \StringTok{"movieId"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{as.data.frame}\NormalTok{()}
\FunctionTok{summary}\NormalTok{(test\_set)}

\NormalTok{train\_set }\OtherTok{\textless{}{-}} \FunctionTok{mutate}\NormalTok{(train\_set, }\AttributeTok{userId =} \FunctionTok{factor}\NormalTok{(userId), }\AttributeTok{movieId =} \FunctionTok{factor}\NormalTok{(movieId))}
\FunctionTok{summary}\NormalTok{(train\_set)}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-39-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-46-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-48-1.pdf}

>>>>

@@

\begin{noteblock}
All the source code of the R-scripts is available on the project's
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/tree/main/r/src}{GitHub
repository}\autocite{edx_capstone_movielens}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{make\_source\_datasets}\NormalTok{()}
\FunctionTok{init\_source\_datasets}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-30-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-31-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-37-1.pdf}
Now we can determine the minimal \emph{RMSE}:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-43-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-44-1.pdf}
Below are worst and best ratings categories:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-46-1.pdf}
